BERT Model Compression and Weight Sharing

### 2019-05-11 paper/paper.tex
Adding model section and resolving conflicts

\newcommand{\bert}{\operatorname{BERT}}
\newcommand{\distortion}{\operatorname{Distortion}}
\newcommand{\prune}{\operatorname{Prune}}
\newcommand{\diffprune}{\operatorname{DiffPrune}}

\newcommand{\btheta}{{\bm{\theta}}}
\newcommand{\bphi}{{\bm{\phi}}}
\newcommand{\bpsi}{{\bm{\psi}}}

@@

>>>>

\newcommand{\bert}{\operatorname{BERT}}
\newcommand{\distortion}{\operatorname{Distortion}}
\newcommand{\prune}{\operatorname{Prune}}
\newcommand{\diffprune}{\operatorname{DiffPrune}}
\newcommand{\rankreduce}{\operatorname{RankReduce}}

\newcommand{\btheta}{{\bm{\theta}}}
\newcommand{\bphi}{{\bm{\phi}}}
\newcommand{\bpsi}{{\bm{\psi}}}
\newcommand{\bvartheta}{{\bm{\vartheta}}}

@@

Given training data $(\bm x_i, y_i)_{i=1}^N$, where $\bm x_i$ are one-hot
encoded tokens, we train by solving the standard maximum likelihood problem
where \[
\min_{\btheta} \sum_{i=1}^N L\pr{y_i, F_\btheta(\bm x_i)}
\]
for some loss function specific to the task. For instance, $L$ is cross-entropy
loss for a (multiclass) classification problem, and $L$ is mean-squared-error
for a regression problem, which correspond to log-likelihoods in Categorical and
Gaussian models, respectively. The innovation of DeepTwist-based training is
perturbing the usual gradient-based optimization by repeatedly distorting the
current weights to achieve compression objectives.

Specifically, the DeepTwist framework presented in \Cref{sec:training} applies a
distortion
step for fixed intervals during training, where \[
\btheta \gets \distortion(\btheta). 
\]
In this section, we discuss a few methods for parameterizing $\distortion$.
Generally, we may imagine a few special cases of distortion, we call a
particular distortion \emph{distortion by layer} if it is applied per layer,
i.e. \begin{align*}
\distortion'(\btheta) = (&\distortion(\bphi_1),\ldots,\distortion(\bphi_n),\\
&\distortion(\bpsi_1),\ldots, \distortion(\bpsi_m)).
\end{align*}
We call a particular distortion \emph{pretrained-only} if \[
\distortion'(\btheta) = \distortion(\bphi).
\]

We list a few methods of parameterizing the distortion step. 
\paragraphdot{Pruning} Consider a parameter vector $\bvartheta$, let
  $t_p$ be the $p$-th quantile of the vector $\bvartheta$. Then \[
  \prune(\bvartheta; p) = \one(|\bvartheta| > t_p) \odot \bvartheta
  \]
  be the vector where values in $\bvartheta$ below $t_p$ are zeroed.

  Pruning is an extremely simple procedure. When $p$ is large, say, $p=0.99$, we
  may take advantage of sparsity of the resulting model by saving locations of
  nonzero entries, and thus reducing the model size in memory. However, such
  compression does not come without potential sacrifices to computational speed,
  as computation needed to be carried out on sparse data structures, which may
  not effectively make use of standard hardware.

  Pruning is an out-of-the-box procedure that works with any models. However, as
  a price of the generality, pruning fails to be tailored to the nature of BERT
  being a pre-trained model. Suppose the pretrained weights $\bphi_0 $ capture
  certain structure of language, then dramatic distortions to $\bphi_0$
  essentially discards the information, especially early in the training loop.
  We may partially remedy this issue by either using a dynamic pruning rate,
  where pruning rate is small early in the training loop or by using
  layer-specific pruning rates, with layers in $\bert$ having a lower pruning
  rate than layers in $G$.

  A direct approach to taking advantage of pretrained models, however, is 
  \emph{difference pruning}, where we ensure that
  the difference between the model and a pretrained model is sparse.
  Effectively, difference pruning ensures a form of weight sharing between
  various BERT-based models for different tasks.

\paragraphdot{Difference Pruning} Consider some pretrained weights
$\bvartheta_0$, define \[
\diffprune(\bvartheta; \bvartheta_0, p ) = \bvartheta_0 + \prune(\bvartheta -
\bvartheta_0 ; p). 
\]
Observe that difference pruning provides no compression benefits when there is
only one fine-tuning task; however, when the number of potential tasks is large,
we only need to save $\btheta_0$ and the sparse matrix $\btheta - \btheta_0,$
and $\btheta_0$ is shared across tasks.

Observe that we can transform any reduction technique by applying the reduction
technique on the difference, and thus the \emph{difference} should be viewed 
(and is implemented) as a decorator to a conventional distortion technique.

\paragraph{Rank Reduction} If $\bvartheta$ is a matrix, then we may apply rank
reduction via singular value decomposition \[
\bvartheta = U\Sigma V^\top
\]
where $\Sigma$ is a diagonal matrix, where $U \in $

We distort via \[
\rankreduce(\bvartheta) = \tilde U \tilde \Sigma \tilde V^\top,
\]
where $\tilde U, \tilde \Sigma, \tilde V$ are conforming matrices such that the
smallest values on the spectrum $\Sigma$ is zeroed out, as well as corresponding
values in $U,V$. If we keep only the $r$ largest spectral values, then we store
$\tilde U, \tilde \Sigma$, an $n\times r$ matrix
